The promise of community driven preprints in ecology and evolution

### 2024-04-30 ms/ms.qmd
Update to figs

*EcoEvoRxiv* is run by the Society for Open, Reliable, and Transparent Ecology and Evolution (SORTEE). Originally launched in 2018 on the Center for Open Science preprint platforms, *EcoEvoRxiv* has grown to become a popular preprint server for researchers in ecology and evolution. It has since been adopted by the California Digital Library (CDL) to be hosted in perpetuity without fees. Editors are researchers across the globe in the fields of ecology and evolution who volunteer their time to screen papers and push new initiatives in the preprint space. *EcoEvoRxiv* accepts both preprints and postprints. Postprints are articles that are already published but are the final version of the published paper before typesetting for the research journal. Postprints are a useful way to ensure published articles are openly accessible and can be posted anytime after publication (assuming the journal has a preprint policy that recognises preprints - though most do). To better understand preprint (and postprint) use on *EcoEvoRxiv*, we downloaded metadata on the articles available on *EcoEvoRxiv* as of `r gsub("UTC", "", dates[2])`. For the purpose of summarising articles on *EcoEvoRxiv* we include both preprints and postprints as 'articles' in our overview analysis.

```{r, fig-summary}
#| label: fig-summary
#| fig-cap: Summary of articles posted to *EcoEvoRxiv*. A) Number of articles (preprints and postprints) published monthly to *EcoEvoRxiv* between 2018 and 2023; B) Countries that are posting articles (preprints and postprints) to *EcoEvoRxiv* as inferred from the country of affiliation of the submitting author; C) Taxa used / covered in the articles posted to *EcoEvoRxiv*; D) Types of articles accepted on *EcoEvoRxiv*. E) Career stage of authors posting to *EcoEvoRxiv* along with the gender of the submitting author.
#| echo: false
#| warning: false
#| message: false

@@

  dates_preprints <- range(preprints$preprint_published_date)
##---------------------##
# Articles on EcoEvoRxiv over time
##---------------------##

  p1 <- ggplot(data2, aes(x = preprint_pub_date)) +
    geom_histogram(fill = "black", color = "black") +
    labs(x = "Published Date", y = "Number of Papers") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + theme(plot.tag = element_text(size = 16, face = "bold"), plot.tag.location = "margin")

##---------------------##
# Countries using EcoEvoRxiv
##---------------------##
  # Create a global map with countried highlighted and shaded according to the frequency of preprints from that country
  # First, we need to get the country names and the number of preprints from each country
    countries <- data2  %>% tabyl(submitting_author_country)  %>% rename("country" = submitting_author_country, "n" = n)  %>% mutate(country = ifelse(country == "NA", "Unknown", country))  %>% arrange(desc(n))  %>% mutate(country=str_to_title(country))

@@

layout <- "
##AAAA##
BBBBBBBB
BBBBBBBB
BBBBBBBB
CCCCDDDD
CCCCDDDD
##EEEE##
##EEEE##
"

@@

p3_pub <- preprints  %>% ggplot(aes(x = oa_status, fill = oa_status)) + geom_bar() + labs(x = "Open Access Status", y = "Number of Papers") + theme_classic() + theme(axis.title = element_text(size = 14)) + scale_fill_manual(values = c("bronze" = "#9b6023", "gold" = "#ffd700", "green" = "#008000", "hybrid" = "#6fa8dc", "NA" = "grey", "closed" = "black"))

>>>>

*EcoEvoRxiv* is run by the Society for Open, Reliable, and Transparent Ecology and Evolution (SORTEE). Originally launched in 2018 on the Center for Open Science preprint platform, *EcoEvoRxiv* has grown to become a popular preprint server for researchers in ecology and evolution. It has since been adopted by the California Digital Library (CDL) to be hosted in perpetuity without fees. Editors are researchers across the globe in the fields of ecology and evolution who volunteer their time to screen papers and push new initiatives in the preprint space. *EcoEvoRxiv* accepts both preprints and postprints. Postprints are articles that are already published but are the final version of the published paper before typesetting for the research journal. Postprints are a useful way to ensure published articles are openly accessible and can be posted anytime after publication (assuming the journal has a preprint policy that recognises preprints - though most do). To better understand preprint (and postprint) use on *EcoEvoRxiv*, we downloaded metadata on the articles available on *EcoEvoRxiv* as of `r gsub("UTC", "", dates[2])`. For the purpose of summarising articles on *EcoEvoRxiv* we include both preprints and postprints as 'articles' in our overview analysis.

```{r, fig-summary}
#| label: fig-summary
#| fig-cap: Summary of articles posted to *EcoEvoRxiv*. A) Number of articles (preprints and postprints) published monthly to *EcoEvoRxiv* between 2018 and 2023. *EcoEvoRxiv* was established in June 2018 before the soft launch in November 2018; B) Countries that are posting articles (preprints and postprints) to *EcoEvoRxiv* as inferred from the country of affiliation of the submitting author; C) Taxa used / covered in the articles posted to *EcoEvoRxiv*; D) Types of articles accepted on *EcoEvoRxiv*. E) Career stage of authors posting to *EcoEvoRxiv* along with the gender of the submitting author.
#| echo: false
#| warning: false
#| message: false

@@

  dates_preprints <- range(preprints$preprint_published_date)
##---------------------##
# Articles on EcoEvoRxiv over time
##---------------------##
  angle = 45
  p1 <- ggplot(data2, aes(x = preprint_pub_date)) +
    geom_histogram(fill = "black", color = "black") +
    labs(x = "Published Date", y = "Number of Papers") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + theme(plot.tag = element_text(size = 16, face = "bold"), plot.tag.location = "margin") + ylim(0, 130) + 
      # Soft Launch of EcoEvoRxiv
    annotate( "segment", x = as.POSIXct("2018-07-01"),  xend = as.POSIXct("2018-07-01"), y = 0, yend = 82, color = "#ffd966", linetype = "dashed", size = 1.5) + 
    annotate("text", x = as.POSIXct("2018-07-01"), y = 85, label = "Soft Launch", color = "#ffd966", size = 4, angle = angle, hjust = 0) +
    # Launch of EcoEvoRxiv
    annotate( "segment", x = as.POSIXct("2018-11-11"),  xend = as.POSIXct("2018-11-11"), y = 0, yend = 82, color = "#2986cc", linetype = "dashed", size = 1.5) + 
    annotate("text", x = as.POSIXct("2018-11-11"), y = 85, label = "Launch", color = "#2986cc", size = 4, angle = angle, hjust = 0) +
    # Spanish and Port
    annotate( "segment", x = as.POSIXct("2023-02-14"),  xend = as.POSIXct("2023-02-14"), y = 0, yend = 82, color = "#3d85c6", linetype = "dashed", size = 1.5) + 
    annotate("text", x = as.POSIXct("2022-11-14"), y = 85, label = "Spanish and Portug.", color = "#3d85c6", size = 4, angle = angle, hjust = 0) +
    # Transition to CDL
    annotate( "segment", x = as.POSIXct("2022-09-14"),  xend = as.POSIXct("2022-09-14"), y = 0, yend = 82, color = "#008000", linetype = "dashed", size = 1.5) + 
    annotate("text", x = as.POSIXct("2022-07-14"), y = 85, label = "CDL", color = "#008000", size = 4, angle = angle, hjust = 0) +
   # IUCN Red Eco submission 29 August 2023
   # 1000 preprints
    annotate( "segment", x = as.POSIXct("2023-08-29"),  xend = as.POSIXct("2023-08-29"), y = 0, yend = 82, color = "#cc0000", linetype = "dashed", size = 1.5) + 
    annotate("text", x = as.POSIXct("2023-04-29"), y = 85, label = "IUCN Report", color = "#cc0000", size = 4, angle = angle, hjust = 0) 
##---------------------##
# Countries using EcoEvoRxiv
##---------------------##
  # Create a global map with countried highlighted and shaded according to the frequency of preprints from that country
  # First, we need to get the country names and the number of preprints from each country
    countries <- data2  %>% tabyl(submitting_author_country)  %>% rename("country" = submitting_author_country, "n" = n)  %>% mutate(country = ifelse(country == "NA", "Unknown", country))  %>% arrange(desc(n))  %>% mutate(country=str_to_title(country))

@@

layout <- "
#AAAAAA#
#AAAAAA#
#AAAAAA#
BBBBBBBB
BBBBBBBB
BBBBBBBB
CCCCDDDD
CCCCDDDD
##EEEE##
##EEEE##
"

@@

p3_pub <- preprints  %>% ggplot(aes(x = oa_status, fill = oa_status)) + geom_bar() + labs(x = "Open Access Status", y = "Number of Papers") + theme_classic() + theme(axis.title = element_text(size = 14), legend.position = "none") + scale_fill_manual(values = c("bronze" = "#9b6023", "gold" = "#ffd700", "green" = "#008000", "hybrid" = "#6fa8dc", "NA" = "grey", "closed" = "black"))
